Comparing and Combining Biomarkers as Principle Surrogates for Time-to-Event Clinical Endpoints

### 2014-06-03 Compare_053014_statsmed_revisions.tex
One more typo found

We develop summary statistics for the comparison of candidate PS for a continuous time-to-event end point. Most existing methods of evaluating candidate PS are not helpful for ranking candidate PS . Often evaluation criteria simply provide a way of testing the binary quality of being or not being a PS, rather than providing a quantitative measurement of surrogate quality. We develop two extensions of the standardized total gain \citet{Huang11}, both of which provide a single values measure of surrogate quality that can be used to compare candidate surrogates for the same clinical output in the same trial. The time-dependent standardized total gain, ($STG(t)$), allows for a single value comparison of marginal CEP models as well as inference for comparison of candidate PS that summaries all information and properly accounts for censoring at and given time point $t$. Over a range of time points, a display of $STG(t)$ versus $t$ curve provides information about surrogate quality waning, as well as a means of comparison of candidate surrogates over time. The time integrated  standardized total gain, ($widetilde{STG}$), which is the time-dependent standardized total gain integrated over the estimated survive times, provides a single value summary of surrogate quality independent of time.

>>>>

We develop summary statistics for the comparison of candidate PS for a continuous time-to-event end point. Most existing methods of evaluating candidate PS are not helpful for ranking candidate PS . Often evaluation criteria simply provide a way of testing the binary quality of being or not being a PS, rather than providing a quantitative measurement of surrogate quality. We develop two extensions of the standardized total gain \citet{Huang11}, both of which provide a single values measure of surrogate quality that can be used to compare candidate surrogates for the same clinical output in the same trial. The time-dependent standardized total gain, ($STG(t)$), allows for a single value comparison of marginal CEP models as well as inference for comparison of candidate PS that summaries all information and properly accounts for censoring at and given time point $t$. Over a range of time points, a display of $STG(t)$ versus $t$ curve provides information about surrogate quality waning, as well as a means of comparison of candidate surrogates over time. The time integrated  standardized total gain, ($\widetilde{STG}$), which is the time-dependent standardized total gain integrated over the estimated distribution of survival times, provides a single value summary of surrogate quality independent of time.
